Towards a Wide-Coverage Grammar

### 2013-06-24 slides/presentation.tex
Pointless -> Ineffective
Final sitdown with Maxime over the slides

\definecolor{darkgreen}{rgb}{0., 0.6, 0.}
\definecolor{darkgold}{rgb}{8., 0.6, 0.}

@@

  We want to build discourse representations in a compositional manner
  by using ACGs. Therefore, we need a wide-coverage ACG grammar.

  Requirements for building a wide-coverage grammar:
    \begin{itemize}
    \item Having linguistic resources to build upon
      \begin{itemize}
      \item Frilex, a formalism-independent lexical database
      \item Frigram, a French interaction grammar
      \end{itemize}
    \item Debugging the grammar
      \begin{itemize}
        \item A toolkit for formulating ACGs as simple metagrammars, for
          examining how they work and for checking their consistency.
      \end{itemize}
    \item Using a formalism that allows for a simple grammar
      \begin{itemize}
      \item The need to express multiple independent constraints led us
        to graphical abstract categorial grammars.
      \end{itemize}
    \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item ACGs are based on types (IILL formulas).
  \item IGs are based on polarized tree descriptions.
  \item Perrier has linked the two.

    \vspace{2 mm}
    \begin{columns}[c]
      \begin{column}{0.1\textwidth}\end{column}
      \begin{column}{0.4\textwidth}
        $$(NP \limp S) \limp N \limp N$$
        $$N \limp (NP \limp S) \limp N$$
      \end{column}
      \begin{column}{0.1\textwidth}$\to$\end{column}
      \begin{column}{0.2\textwidth}
        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
      \end{column}
      \begin{column}{0.25\textwidth}\end{column}
    \end{columns}
  \item The IG of Frigram use different constraints in tree
    descriptions.
    \begin{itemize}
    \item The deductive syntactic model of IGs is built on a wider range
      of linear logic primitives ($\otimes$, $\&$).
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \begin{block}{Abstract languages}
    Languages of (abstract) terms constrained by the constants they are
    built on and their types. Defined by a \textcolor{red}{signature}
    and a \textcolor{darkgreen}{type}.

@@

  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Two different lexicons interpret the same abstract terms.
      \item Instead of referring to the two ACGs $\mathopen{<}
        \Sigma_{Synt}, \Sigma_{String}, \mathcal{L}_{syntax}, S
        \mathclose{>}$ and $\mathopen{<} \Sigma_{Synt}, \Sigma_{Sem},
        \mathcal{L}_{sem}, S \mathclose{>}$, we refer to this single
        diagram.
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/double-acg.pdf}
    \end{column}
  \end{columns}
\end{frame}

@@

  In order to stay in the scope of ACGs, our diagrams must be shaped as
  arborescences (at most one parent per node).

@@

      $\implies$ Complex.
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/fail1.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.25\textwidth}
      {\small Syntactic constraints deal with the syntax-semantics
        interface.}

      $\implies$ Complex.
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/fail2.pdf}
      \end{center}
    \end{column}
  \end{columns}
\end{frame}

@@

  We will need to be able to specify constraints independently if we are
  to write a manageably simple wide-coverage grammar.

@@

      $\implies$ Pointless.
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-1.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.25\textwidth}
      Third constraint reimplements first and second constraint.

      $\implies$ Pointless.
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-2.pdf}
      \end{center}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Composing ACG Patterns III}

  Therefore, we would like to able to do something like this (i.e., not
  an arborescence).

  \begin{columns}[c]
    \begin{column}{0.6\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/final.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.4\textwidth}
      How do we assign languages to nodes of the graph? \\
      $\implies$ Graphical Abstract Categorial Grammars (G-ACGs)

      \vspace{5 mm}

      Formally, we have:
        \begin{itemize}
        \item A DAG.
        \item Nodes labelled with:
          \begin{itemize}
          \item Signatures.
          \item Distinguished types.
          \end{itemize}
        \item Edges labelled with:
          \begin{itemize}
          \item Lexicons.
          \end{itemize}
        \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\section{Incompatible Intuitions in G-ACGs}

@@

  \begin{block}{Algebraic Perspective}
    \begin{itemize}
    \item Every node is a language.
    \item An arrow means that a language is produced from another by
      mapping through a lexicon.
    \item Two arrows converging on a node mean an intersection of
      languages.
    \end{itemize}
  \end{block}
  \begin{block}{Generative Perspective}
    \begin{itemize}
    \item Every node is a term.
    \item An arrow means that a term is produced from another by
      applying a lexicon.
    \item Two arrows converging on a node mean that the two (abstract)
      terms yield the same (object) term.
    \end{itemize}
  \end{block}
\end{frame}

@@

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item Every node is a language.
      \item An arrow means that a language is produced from another by
        mapping through a \textcolor{darkgreen}{lexicon}.
      \item Two arrows converging on a node mean an
        \textcolor{red}{intersection} of languages.
      \end{itemize}

@@

  \begin{columns}[t]
    \begin{column}{0.5\textwidth}
      \includegraphics[height=0.6\textheight]{../diagrams/both-patterns.pdf}

      How do we get the languages of meanings expressible only by
      syntactically correct sentences?
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[height=0.6\textheight]{../diagrams/diamond-grammar.pdf}

@@

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item \textcolor{red}{Every node is a term.}
      \item An arrow means that a term is produced from another by
        applying a \textcolor{darkgreen}{lexicon}.
      \item Two arrows converging on a node mean that the two (abstract)
        terms yield \textcolor{red}{the same} (object) term.
      \end{itemize}

@@

\begin{frame}
  \frametitle{Only Different Points of View?}

  \begin{itemize}
  \item $\forall \, \mathcal{G}, u$ where $\mathcal{G}$ is arborescent,
    we have $\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)
    = \mathcal{O}_{\mathcal{G}}(u)$.
  \item However, $\exists \, \mathcal{G}, u$ such that
    $\mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$.
  \item The two different ``perspectives'' lead to different languages.
    $\implies$ They are not interchangeable points of view.
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item The three language families we defined above are\ldots
    \begin{itemize}
    \item {\ldots}increasingly constraining, $\forall \, \mathcal{G},
      u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
      \mathcal{E}_{\mathcal{G}}(u) \supseteq
      \mathcal{P}_{\mathcal{G}}(u)$.
    \item {\ldots}increasingly expressive, $\mathcal{I} \subseteq
      \mathcal{E} \subseteq \mathcal{P}$.
    \end{itemize}
  \item Intrinsic languages are just the abstract languages of ACGs.
    \begin{itemize}
    \item $\mathcal{I} = \mathcal{A}$
    \end{itemize}
  \item Extrinsic languages are simple extensions of ACGs that add
    intersections.
    \begin{itemize}
    \item $\mathcal{E} = \mathcal{O}^{\mathcal{L}{\cap}}$
    \end{itemize}
  \item Pangraphical languages are faithful to the generative
    perspective and subsume the previous classes of languages.
    \begin{itemize}
    \item They solve both the example problems we saw before
      (\emph{meanings of grammatical sentences} and the \emph{diamond}).
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Direct embedding of IGs into ACGs based on Perrier's theorem
    conjectured not to be practical.
  \item Created a toolkit to serve as basis for more explorations or
    implementations.
  \item Formalized graphical abstract categorial grammars.
    \begin{itemize}
    \item Helps solve multiple constraints elegantly.
    \item Facilitates discussion and inquiry into compositions of ACGs.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Make a closer examination of the relationship between extrinsic
    and pangraphical languages in G-ACGs.
  \item Determine whether ACG object languages are closed on intersection
    or not.
  \item Build a wide-coverage G-ACG and study the patterns used in the
    graphical architecture.
  \item Try to express the multiple independent polarized features of IGs
    using G-ACGs.
  \item Assess the applicability of G-ACG architectures at other levels
    of description, e.g. discourse structure.
  \end{itemize}
\end{frame}

>>>>

\definecolor{darkgreen}{rgb}{0., 0.6, 0.}
\definecolor{darkgold}{rgb}{8., 0.6, 0.}
\definecolor{limegreen}{RGB}{175, 255, 175}
\definecolor{limeblue}{RGB}{200, 200, 235}

@@

  \textbf{Objective:} Build discourse representations using ACGs: \\a
  wide-coverage ACG grammar.

  \vfill

  \begin{itemize}
  \item Having linguistic resources to build upon
    \begin{itemize}
    \item Frilex, a formalism-independent lexical database
    \item Frigram, a French interaction grammar
    \end{itemize}
    \vfill
  \item Debugging the grammar
    \begin{itemize}
    \item A toolkit for formulating ACGs as simple metagrammars, for
      examining how they work and for checking their consistency.
    \end{itemize}
    \vfill
  \item Using a formalism that allows for a simple grammar
    \begin{itemize}
    \item Express multiple independent constraints \\ $\implies$ graphical abstract categorial grammars.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item ACGs are based on types (IILL formulas).
  \item IGs are based on polarized tree descriptions.
    \vfill
  \item[$\Rightarrow$] Perrier has linked the two.

    \vspace{2 mm}
    \begin{columns}[c]
      \begin{column}{0.1\textwidth}\end{column}
      \begin{column}{0.4\textwidth}
        $$(NP \limp S) \limp N \limp N$$
        $$N \limp (NP \limp S) \limp N$$
      \end{column}
      \begin{column}{0.1\textwidth}$\to$\end{column}
      \begin{column}{0.2\textwidth}
        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
      \end{column}
      \begin{column}{0.25\textwidth}\end{column}
    \end{columns}
    \vfill
  \item IG of Frigram use different constraints in tree descriptions.
    \begin{itemize}
    \item The deductive syntactic model of IGs is built on a wider range
      of linear logic primitives ($\otimes$, $\&$).
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \begin{block}{Abstract languages}
    %% Languages of (abstract) terms constrained by the constants they are
    %% built on and their types. Defined by a \textcolor{red}{signature}
    %% and a \textcolor{darkgreen}{type}.

    Lambda terms constrained by the types of constants. \\ Defined by a
    \textcolor{red}{signature} and a \textcolor{darkgreen}{type}.

@@

  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Two different lexicons interpret the same abstract terms.
      \item Instead of referring to the two ACGs $\mathopen{<}
        \Sigma_{Synt}, \Sigma_{String}, \mathcal{L}_{syntax}, S
        \mathclose{>}$ and $\mathopen{<} \Sigma_{Synt}, \Sigma_{Sem},
        \mathcal{L}_{sem}, S \mathclose{>}$, refer to this single
        diagram.
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/double-acg.pdf}
    \end{column}
  \end{columns}
\end{frame}

@@

  ACG diagrams are arborescences.

@@

      \onslide<2->{$\implies$ Complex.}
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/fail1.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.25\textwidth}
      {\small Syntactic constraints deal with the syntax-semantics
        interface.}

      \onslide<3->{$\implies$ Complex.}
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/fail2.pdf}
      \end{center}
    \end{column}
  \end{columns}
\end{frame}

@@

  Writing a simple wide-coverage grammar implies specifying constraints
  independently.

@@

      \onslide<2->{$\implies$ Ineffective.}
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-1.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.25\textwidth}
      Third constraint reimplements first and second constraint.

      \onslide<3->{$\implies$ Ineffective.}
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-2.pdf}
      \end{center}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Graphical Abstract Categorial Grammars (G-ACGs)}

  \begin{columns}[c]
    \begin{column}{0.6\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/final.pdf}
      \end{center}
    \end{column}

    \pause

    \begin{column}{0.4\textwidth}
        \begin{itemize}
        \item A DAG.
        \item Nodes labelled with:
          \begin{itemize}
          \item Signatures.
          \item Distinguished types.
          \end{itemize}
        \item Edges labelled with:
          \begin{itemize}
          \item Lexicons.
          \end{itemize}
        \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\section{Semantics of G-ACGs Nodes}

@@

  \setbeamercolor{block title}{bg=limeblue}
  \setbeamercolor{block body}{bg=limeblue}
  \begin{block}{Algebraic Perspective}
    \begin{itemize}
    \item Nodes as \emph{languages}.
    \item Edges define \emph{languages} in terms of other
      \emph{languages} using lexicons and intersections.
    \end{itemize}
  \end{block}
  \setbeamercolor{block title}{bg=limegreen}
  \setbeamercolor{block body}{bg=limegreen}
  \begin{block}{Generative Perspective}
    \begin{itemize}
    \item Nodes as \emph{terms}.
    \item Edges describe how \emph{terms} are obtained from other
      \emph{terms} by applying lexicons.
    \end{itemize}
  \end{block}
\end{frame}

@@

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \setbeamercolor{block title}{bg=limeblue}
      \setbeamercolor{block body}{bg=limeblue}
      \begin{block}{}
        \begin{itemize}
        \item Every node is a language.
        \item An arrow means that a language is produced from another by
          mapping through a \textcolor{darkgreen}{lexicon}.
        \item Two arrows converging on a node mean an
          \textcolor{red}{intersection} of languages.
        \end{itemize}
      \end{block}

@@

  \begin{columns}[t]
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.6\textheight]{../diagrams/both-patterns.pdf}
      \end{center}

      How do we get the language of meanings expressible only by
      syntactically correct sentences?
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.6\textheight]{../diagrams/diamond-grammar.pdf}
      \end{center}

@@

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \setbeamercolor{block title}{bg=limegreen}
      \setbeamercolor{block body}{bg=limegreen}
      \begin{block}{}
        \begin{itemize}
        \item Every node is a \textcolor{red}{term}.
        \item An arrow means that a term is produced from another by
          applying a \textcolor{darkgreen}{lexicon}.
        \item Two arrows converging on a node mean that the two (abstract)
          terms yield \textcolor{red}{the same} (object) term.
        \end{itemize}
      \end{block}

@@

\begin{frame}
  \frametitle{Algebraic and Generative Relations}

  \begin{itemize}
  \item $\forall \, \mathcal{G}, u$ where $\mathcal{G}$ is arborescent:
    $$\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)
    = \mathcal{O}_{\mathcal{G}}(u)$$
    \pause
  \item However, $\exists \, \mathcal{G}, u$ such that
    $$\mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$$
    \vfill
    \pause
  \item[$\Rightarrow$] They are not interchangeable points of view.
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item The three language families we defined above are\ldots
    \begin{itemize}
    \item {\ldots}increasingly constraining, $\forall \, \mathcal{G},
      u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
      \mathcal{E}_{\mathcal{G}}(u) \supseteq
      \mathcal{P}_{\mathcal{G}}(u)$.
    \item {\ldots}increasingly expressive, $\mathcal{I} \subseteq
      \mathcal{E} \subseteq \mathcal{P}$.
    \end{itemize}
  \vfill
  \item Intrinsic languages are just the abstract languages of ACGs.
    \begin{itemize}
    \item $\mathcal{I} = \mathcal{A}$
    \end{itemize}
  \vfill
  \item Extrinsic languages are simple extensions of ACGs that add
    intersections.
    \begin{itemize}
    \item $\mathcal{E} = \mathcal{O}^{\mathcal{L}{\cap}}$
    \end{itemize}
  \vfill
  \item Pangraphical languages are faithful to the generative
    perspective and subsume the previous classes of languages.
    \begin{itemize}
    \item They solve both the example problems we saw before
      (\emph{meanings of grammatical sentences} and the \emph{diamond}).
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Direct embedding of IGs into ACGs based on Perrier's theorem
    conjectured not to be practical.
  \vfill
  \item Created a toolkit to serve as basis for more explorations or
    implementations.
  \vfill
  \item Formalized graphical abstract categorial grammars.
    \begin{itemize}
    \item Helps solve multiple constraints elegantly.
    \item Facilitates discussion and inquiry into compositions of ACGs.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Are extrinsic languages as expressive as pangraphical
    ones? $$\mathcal{P} \subseteq \mathcal{E}$$
  \item Are ACG object languages closed on
    intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
  \item Express the multiple polarized features of IGs in G-ACGs.
    \vfill
  \item Build and study a wide-coverage G-ACG.
    \vfill
  \item Apply G-ACGs to discourse.
  \end{itemize}
\end{frame}
